---
description: |
  "This is a child document for README.Rmd to isolate What's New updates (last section of README)"
  "see README.Rmd's last chunk."
  "No need to knit this. Knitting README will automatically include this."
---

```{r setup-news, include=FALSE}
library(knitr)
opts_chunk$set( echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE)
```


```{r, echo=FALSE, eval=FALSE, message='hide', fig.asp=1, out.width="50%"}
## template chunk - don't change/delete
```

## What's New

#### May 2026

* bug fix implemented for [#377](https://github.com/mixOmicsTeam/mixOmics/issues/377) fixed vector `xlim` handling in the `graphics` branch of `plotLoadings()` contribution plots
* enhancement request [#381](https://github.com/mixOmicsTeam/mixOmics/issues/381) expanded test coverage for model utilities and internal helper paths, with small fixes made for issues exposed by the new tests

#### April 2026
** Version 6.36.0 **

[Bioconductor release version 6.36.0](https://www.bioconductor.org/packages/release/bioc/html/mixOmics.html) released 29th April 2026
mixOmics can now be downloaded using [Bioconductor version 3.23](https://bioconductor.org/news/bioc_3_23_release/) and is compatible with R 4.6.0

* bug fix implemented for [#374](https://github.com/mixOmicsTeam/mixOmics/issues/374) replaced deprecated `aes_string()` with `aes(.data[[...]])` in plotting functions
* bug fix implemented for [#379](https://github.com/mixOmicsTeam/mixOmics/issues/379) replaced deprecated `size` aesthetic with `linewidth` in ggplot2 line geoms

#### April 2025
** Version 6.32.0 **

[Bioconductor release version 6.32.0](https://www.bioconductor.org/packages/release/bioc/html/mixOmics.html) released 17th April 2025
mixOmics can now be downloaded using [Bioconductor version 3.21](https://bioconductor.org/news/bioc_3_21_release/) and is compatible with R 4.5.0

* feature request [#345](https://github.com/mixOmicsTeam/mixOmics/issues/345) updated functionality for `plotLoadings()` so can plot in ggplot2 style and customise aesthetics
* bug fix implemented for [#357](https://github.com/mixOmicsTeam/mixOmics/issues/357) `plotIndiv()` not handing `pch` ordering correctly
* enhancement request [#332](https://github.com/mixOmicsTeam/mixOmics/issues/332) increased test coverage for `plotIndiv()`
* implemented a new unit testing framework for plotting functions using `vdiffr` package

#### March 2025

* enhancement request [#353](https://github.com/mixOmicsTeam/mixOmics/issues/353) better error message in `perf()` when one sample in a class
* enhancement request [#340](https://github.com/mixOmicsTeam/mixOmics/issues/340) expand test coverage for main functions
* enhancement request [#336](https://github.com/mixOmicsTeam/mixOmics/issues/336) streamline multiblock functions by removing `scheme` and `init` args

Also explored potential unusual behaviour of:
* zero variance handling in `block.splsda` [#352](https://github.com/mixOmicsTeam/mixOmics/issues/352) 
* `perf()` giving non-intuitve per-class error rates [#355](https://github.com/mixOmicsTeam/mixOmics/issues/355)

#### November 2024

* enhancement request [#216](https://github.com/mixOmicsTeam/mixOmics/issues/216) implemented parallel processing using `BPPARAM` across all `tune()` functions
* feature request [#335](https://github.com/mixOmicsTeam/mixOmics/issues/335) added `seed` argument to `perf()` functions for better reproducibility
* feature request [#334](https://github.com/mixOmicsTeam/mixOmics/issues/334) added `seed` argument to `tune()` functions for better reproducibility
* bug fix implemented for [#303](https://github.com/mixOmicsTeam/mixOmics/issues/303) multiple solutions found in `perf()` returns error
* bug fix implemented for [#307](https://github.com/mixOmicsTeam/mixOmics/issues/307) `plotIndiv()` ellipses colours not matching points, now sample group order is respected and colours can be customised for points and ellipses
* updated documentation to fix issue [#297](https://github.com/mixOmicsTeam/mixOmics/issues/297) broken link in bookdown vignette
* updated documentation to fix issue [#296](https://github.com/mixOmicsTeam/mixOmics/issues/296) typo in vignette

The performance assessment and parameter tuning workflow has been streamlined as described in issue [#343](https://github.com/mixOmicsTeam/mixOmics/issues/343)

* New function: `perf.assess()` 
This function essentially runs `perf()` on final model but only returns performance metrics for the number of components used in the final model. 
Designed to be used in the final step of the workflow for quantifying final model performance. Outputs a list of values but no plotting functionality avaliable. 
See [PR #344](https://github.com/mixOmicsTeam/mixOmics/pull/344) for more details. 

* Additional functionality for `tune()` functions and new `tune()` functions created
`tune()` can now be used in its original capacity (to tune number of variables and components simultaneously) or just to tune number of components by internally calling `perf()`. 
Designed to be used for tuning both components and variables to keep across (s)PCA, (s)PLS, (s)PLSDA, block (s)PLSDA and mint (s)PLSDA models
See [PR #348](https://github.com/mixOmicsTeam/mixOmics/pull/348) for more details.  

#### October 2024
** Version 6.30.0 **

[Bioconductor release version 6.30.0](https://bioconductor.org/packages/release/bioc/html/mixOmics.html) released end of October 2024
Minor bug fixes and updated deprecated code and unit tests, no major code changes and no changes to user experience of mixOmics. 

* bug fix implemented for [#293](https://github.com/mixOmicsTeam/mixOmics/issues/293) `splsda()` example code error

#### March 2022

* bug fix implemented for [Issue #196](https://github.com/mixOmicsTeam/mixOmics/issues/196). `perf()` can now handle features with a `(s)pls` which have near zero variance.
* bug fix implemented for [Issue #192](https://github.com/mixOmicsTeam/mixOmics/issues/192). `predict()` can now handle when the testing and training data have their columns in different orders.
* bug fix implemented for [Issue #178](https://github.com/mixOmicsTeam/mixOmics/issues/178). If the `indY` parameter is used in `block.spls()`, `circosPlot()` can now properly identify the $Y$ dataframe.
* bug fix implemented for [Issue #172](https://github.com/mixOmicsTeam/mixOmics/issues/172). `perf()` now returns values for the `choice.ncomp` component when `nrepeat` $< 3$ whereas before it would just return `NA`s.
* bug fix implemented for [Issue #171](https://github.com/mixOmicsTeam/mixOmics/issues/171). `cim()` now can take `pca` objects as input.
* bug fix implemented for [Issue #161](https://github.com/mixOmicsTeam/mixOmics/issues/161). `tune.spca()` can now handle `NA` values appropriately.
* bug fix implemented for [Issue #150](https://github.com/mixOmicsTeam/mixOmics/issues/150). Provided users with a specific error message for when `plotArrow()` is run on a `(mint).(s)plsda` object.
* bug fix implemented for [Issue #122](https://github.com/mixOmicsTeam/mixOmics/issues/122). Provided users with a specific error message for when a `splsda` object that has only one sample associated with a given class is passed to `perf()`.
* bug fix implemented for [Issue #120](https://github.com/mixOmicsTeam/mixOmics/issues/120). `plotLoadings()` now returns the loading values for features from **all** dataframes rather than just the last one when operating on a `(mint).(block).(s)plsda` object.
* bug fix implemented for [Issue #43](https://github.com/mixOmicsTeam/mixOmics/issues/43). Homogenised the way in which `tune.mint.splsda()` and `perf.mint.splsda()` calculate balanced error rate (BER) as there was disparity between them. Also made the global BER a weighted average of BERs across each study.
* enhancement implemented for [Issue #30/#34](https://github.com/mixOmicsTeam/mixOmics/issues/34). The parameter `verbose.call` was added to most of the methods. This parameter allows users to access the specific values input into the call of a function from its output.
* bug fix implemented for [Issue #24](https://github.com/mixOmicsTeam/mixOmics/issues/24). `background.predict()` can now operate on `mint.splsda` objects and can be used as part of `plotIndiv()`.

#### July 2021

* new function `plotMarkers` to visualise the selected features in block analyses (see https://github.com/mixOmicsTeam/mixOmics/issues/134)
* `tune.spls` now able to tune the selected variables on both `X` and `Y`. See `?tune.spls`
* new function `impute.nipals` to impute missing values using the nipals algorithm
* new function `tune.spca` to tune the number of selected variables for pca components
* `circosPlot` now has methods for `block.spls` objects. It can now handle similar feature names across blocks. It is also much more customisable. See advanced arguments in `?circosPlot`
* new `biplot` function for `pca` and `pls` objects. See `?mixOmics::biplot`
* `plotDiablo` now takes `col.per.group` (see #119)

#### April 2020

* weighted consensus plots for DIABLO objects now consider per-component weights

#### March 2020

* `plotIndiv` now supports (weighted) consensus plots for block analyses. See the example in [this issue](https://github.com/mixOmicsTeam/mixOmics/issues/57)
* `plotIndiv(..., ind.names=FALSE)` [warning issue](https://github.com/mixOmicsTeam/mixOmics/issues/59) now fixed

#### January 2020

* `perf.block.splsda` now supports calculation of combined AUC
* `block.splsda` bug which could drop some classes with `near.zero.variance=TRUE` now fixed
